multiset_hyper2

### 2022-08-01 inst/multiset_hyper2.Rmd
more exacting test

```{r defineM}
M <- apply(multiset(c(1,1,2,2,3)),2,function(x){letters[x]})
noquote(M)
```

```{r definetfun}
tfun <- function(help,strength){
  out <- 0
  for(i in seq_len(ncol(M))){
     vec <- M[,i]
     out <- out +
       loglik(strength,cheering3(v=vec,e=c(a=1,b=2,c=3),help=help),log=FALSE)
  }
  return(out)
}
```

```{r showtisconstant,cache=TRUE}
print(tfun(help=(1:3)/10, strength= c(a=0.1,b=0.3,c=0.6)))
print(tfun(help=(7:5)/22, strength= c(a=0.3,b=0.2,c=0.5)))
```

@@

Above we see that $\mathbf{do}$ and $\mathbf{re}$ come first and last
(or last and first), and $\mathbf{me}$ and $\mathbf{fa}$ come third
and fourth (or fourth and third), while $\mathbf{sol}$ comes second.
All four possibilities have the same probability and all four are
admissible order statistics for the five solfeggio syllables.

@@

>>>>

```{r defineM}
M <- apply(multiset(c(1,1,2,2,3)),2,function(x){letters[x]})
colnames(M) <- rep("",ncol(M))   # tidier and more compact appearance
M <- noquote(M)
M
```

Above, we use `multiset()` to enumerate the sample space of five
competitors comprising two twins denoted `a`, two twins denoted `b`,
and a singleton `c`.  This document discusses the issue of whether the
twins are identical twins and hence indistinguishable.  Matrix `M` has
${5\choose2\,2\,1}=30$ columns, one per possible result.  We will
calculate the likelihood of each column and calculate their sum:

```{r definetfun}
tfun <- function(help,strength,M){
  out <- 0
  for(i in seq_len(ncol(M))){
     vec <- M[,i]
     out <- out +
       loglik(strength,cheering3(v=vec,e=c(a=1,b=2,c=3),help=help),log=FALSE)
  }
  return(out)
}
```

```{r showtisconstant,cache=TRUE}
print(tfun(help=(1:3)/10, strength= c(a=0.1,b=0.3,c=0.6),M))
print(tfun(help=(7:5)/22, strength= c(a=0.3,b=0.2,c=0.5),M))
```

@@

heAbove we see that $\mathbf{do}$ and $\mathbf{re}$ come first and last
(or last and first), and $\mathbf{me}$ and $\mathbf{fa}$ come third
and fourth (or fourth and third), while $\mathbf{sol}$ comes second.
All four possibilities have the same probability and all four are
admissible order statistics for the five solfeggio syllables.

@@

## A more exacting numerical test

```{r,exactingtest,cache=TRUE}
rs <- fillup(rp(100,dirichlet(c(a=1,b=1,c=1)))) # random strengths
colnames(rs) <- letters[1:3]
head(rs)
f <- function(s){
  jj <- s
  names(jj) <- letters[1:3]
  tfun(help=1:3,strength=jj,M=M)
}
x <- apply(rs,1,f)
```

```{r}
hist(x-mean(x))
```

Above we see very tight clustering about the mean value of 0.25
